---
title: "R2BEAT_workflows"
author: "Giulio Barcaroli"
date: "6/10/2021"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Workflow examples with R2BEAT 
# Scenario 1
Only a sampling frame containing the units of the population 
of reference is available, no previous round of the sampling survey to be 
planned


```{r}
# Install last version of R2BEAT and ReGenesees
#install.packages("devtools")
#devtools::install_github("barcaroli/R2BEAT",dependencies=FALSE)
library("R2BEAT")
```

```{r}
packageVersion("R2BEAT")
```


```{r}
## Sampling frame
load("pop.RData")
str(pop)
```

## Precision constraints


```{r}

cv <- as.data.frame(list(DOM=c("DOM1","DOM2"),
                         CV1=c(0.02,0.03),
                         CV2=c(0.03,0.06),
                         CV3=c(0.03,0.06),
                         CV4=c(0.05,0.08)))
cv
```

## Sensitivity analysis


```{r}
sens_min_SSU <- sensitivity_min_SSU (
             samp_frame=pop,
             errors=cv,
             id_PSU="municipality",
             id_SSU="id_ind",
             strata_var="stratum",
             target_vars=c("income_hh","active","inactive","unemployed"),
             deff_var="stratum",
             domain_var="region",
             delta=1,
             deff_sugg=1,
             min=30,
             max=80,
             plot=TRUE)
```


## Preparation of inputs for allocation steps


```{r}
## Preparation of inputs for allocation steps
samp_frame <- pop
samp_frame$one <- 1
id_PSU <- "municipality"  
id_SSU <- "id_ind"        
strata_var <- "stratum"   
target_vars <- c("income_hh","active","inactive","unemployed")   
deff_var <- "stratum"     
domain_var <- "region"  
delta =  1       # households = survey units
minimum <- 50    # minimum number of SSUs to be interviewed in each selected PSU
deff_sugg <- 1.5 # suggestion for the deff value
 
inp1 <- prepareInputToAllocation1(samp_frame,
                                id_PSU,
                                id_SSU,
                                strata_var,
                                target_vars,
                                deff_var,
                                domain_var,
                                minimum,
                                delta,
                                deff_sugg)
```

```{r}
head(inp1$deff)
```

```{r}
head(inp1$effst)
```

```{r}
head(inp1$rho)
```

```{r}
head(inp1$psu_file)
```

```{r}
head(inp1$des_file)
```


## Allocation


```{r}
inp1$desfile$MINIMUM <- 50
alloc1 <- beat.2st(stratif = inp1$strata, 
                  errors = cv, 
                  des_file = inp1$des_file, 
                  psu_file = inp1$psu_file, 
                  rho = inp1$rho, 
                  deft_start = NULL,
                  effst = inp1$effst, 
                  minPSUstrat = 2,
                  minnumstrat = 50
                  )
```


## Selection of PSUs (I stage)


```{r}
set.seed(1234)
sample_1st <- select_PSU(alloc1, type="ALLOC", pps=TRUE)
```

```{r}
sample_1st$PSU_stats
```


## Selection of SSUs (II stage)


```{r}
samp <- select_SSU(df=pop,
                   PSU_code="municipality",
                   SSU_code="id_ind",
                   PSU_sampled=sample_1st$sample_PSU,
                   verbose=TRUE)
```



```{r}
## Plot of weights distribution
par(mfrow=c(1, 2))
boxplot(samp$weight,col="grey")
title("Weights distribution (total sample)",cex.main=0.7)
boxplot(weight ~ region, data=samp,col="grey")
title("Weights distribution by region",cex.main=0.7)
par(mfrow=c(1, 2))
boxplot(weight ~ province, data=samp,col="grey")
title("Weights distribution by province",cex.main=0.7)
boxplot(weight ~ stratum, data=samp,col="grey")
title("Weights distribution by stratum",cex.main=0.7)
```


## Precision constraints compliance control (by simulation)


```{r}
df=pop
df$one <- 1
PSU_code="municipality"
SSU_code="id_ind"
target_vars <- c("income_hh",
                 "active",
                 "inactive",
                 "unemployed")  
```


```{r}
# Domain level = national
domain_var <- "one"
set.seed(1234)
eval11 <- eval_2stage(df,
                    PSU_code,
                    SSU_code,
                    domain_var,
                    target_vars,
                    sample_1st$sample_PSU,
                    nsampl=100, 
                    writeFiles=FALSE,
                    progress=TRUE) 
eval11$coeff_var
```

```{r}
# Domain level = regional
domain_var <- "region"
set.seed(1234)
set.seed(1234)
eval12 <- eval_2stage(df,
                    PSU_code,
                    SSU_code,
                    domain_var,
                    target_vars,
                    sample_1st$sample_PSU,
                    nsampl=100, 
                    writeFiles=FALSE,
                    progress=TRUE) 
eval12$coeff_var
```

```{r}
alloc1$sensitivity
```


```{r}
save(samp,file="sample.RData")
```

# Scenario 2
One previous round of the sampling survey is available.

## Analysis of sampled data


```{r}
#devtools::install_github("DiegoZardetto/ReGenesees",dependencies = FALSE)
library(ReGenesees)
```

```{r}
load("sample.RData")
str(samp)
```

```{r}
## Sample design description
samp$stratum_2 <- as.factor(samp$stratum_2)
sample.des <- e.svydesign(samp, 
                          ids= ~ municipality + id_hh, 
                          strata = ~ stratum_2, 
                          weights = ~ weight,
                          self.rep.str = ~ SR,
                          check.data = TRUE)
```


```{r}
## Find and collapse lonely strata
ls <- find.lon.strata(sample.des)
if (!is.null(ls)) sample.des <- collapse.strata(sample.des)
```


```{r}
## Calibration with known totals
totals <- pop.template(sample.des,
             calmodel = ~ sex : cl_age, 
             partition = ~ region)
totals <- fill.template(pop, totals, mem.frac = 10)
sample.cal <- e.calibrate(sample.des, 
                          totals,
                          calmodel = ~ sex : cl_age, 
                          partition = ~ region,
                          calfun = "logit",
                          bounds = c(0.3, 2.6), 
                          aggregate.stage = 2,
                          force = FALSE)
```

 
## Preparation of inputs for allocation steps


```{r}
samp_frame <- pop
RGdes <- sample.des
RGcal <- sample.cal
strata_var <- c("stratum")      
target_vars <- c("income_hh",
                 "active",
                 "inactive",
                 "unemployed")   
weight_var <- "weight"
deff_var <- "stratum"            
id_PSU <- c("municipality")      
id_SSU <- c("id_hh")             
domain_var <- c("region") 
delta <- 1                   
minimum <- 50                

inp2 <- prepareInputToAllocation2(
        samp_frame,  # sampling frame
        RGdes,       # ReGenesees design object
        RGcal,       # ReGenesees calibrated object
        id_PSU,      # identification variable of PSUs
        id_SSU,      # identification variable of SSUs
        strata_var,  # strata variable
        target_vars, # target variables
        deff_var,    # deff variable
        domain_var,  # domain variable
        delta,       # Average number of SSUs for each selection unit
        minimum      # Minimum number of SSUs to be selected in each PSU
      )
```


```{r}
head(inp2$strata)
```


```{r}
head(inp2$deff)
```


```{r}
head(inp2$effst)
```


```{r}
head(inp2$rho)
```


```{r}
head(inp2$psu_file)
```

```{r}
head(inp2$des_file)
```


## Allocation


```{r}
set.seed(1234)
inp2$des_file$MINIMUM <- 50
alloc2 <- beat.2st(stratif = inp2$strata, 
                  errors = cv, 
                  des_file = inp2$des_file, 
                  psu_file = inp2$psu_file, 
                  rho = inp2$rho, 
                  deft_start = NULL, 
                  effst = inp2$effst,
                  minnumstrat = 2, 
                  minPSUstrat = 2)
```



## Selection of PSUs (I stage)

```{r}
set.seed(1234)
sample_1st <- select_PSU(alloc2, type="ALLOC", pps=TRUE)
```

```{r}
sample_1st$PSU_stats
```


## Selection of SSUs (II stage)


```{r}
set.seed(1234)
samp <- select_SSU(df=pop,
                   PSU_code="municipality",
                   SSU_code="id_ind",
                   PSU_sampled=sample_1st$sample_PSU,
                   verbose=TRUE)
```

```{r}
nrow(samp)
sum(alloc2$alloc$ALLOC[-nrow(alloc2$alloc)])
```


```{r}
nrow(pop)
sum(samp$weight)
```


```{r}
## Plot of weights distribution
par(mfrow=c(1, 2))
boxplot(samp$weight,col="grey")
title("Weights distribution (total sample)",cex.main=0.7)
boxplot(weight ~ region, data=samp,col="grey")
title("Weights distribution by region",cex.main=0.7)
par(mfrow=c(1, 2))
boxplot(weight ~ province, data=samp,col="grey")
title("Weights distribution by province",cex.main=0.7)
boxplot(weight ~ stratum, data=samp,col="grey")
title("Weights distribution by stratum",cex.main=0.7)
```


## Precision constraints compliance control (by simulation)


```{r}
df=pop
df$one <- 1
PSU_code="municipality"
SSU_code="id_ind"
target_vars <- c("income_hh",
                 "active",
                 "inactive",
                 "unemployed")  
```


```{r}
# Domain level = national
domain_var <- "one"
set.seed(1234)
eval21 <- eval_2stage(df,
                    PSU_code,
                    SSU_code,
                    domain_var,
                    target_vars,
                    PSU_sampled=sample_1st$sample_PSU,
                    nsampl=100, 
                    writeFiles=FALSE,
                    progress=TRUE) 
eval21$coeff_var
```


```{r}
# Domain level = regional
domain_var <- "region"
set.seed(1234)
eval22 <- eval_2stage(df,
                    PSU_code,
                    SSU_code,
                    domain_var,
                    target_vars,
                    PSU_sampled=sample_1st$sample_PSU,
                    nsampl=100, 
                    writeFiles=FALSE,
                    progress=TRUE) 
eval22$coeff_var
```


```{r}
alloc2$sensitivity
```

